---
title: "SafetyCulture Headend Example"
author: "Michael Mantfeld"
date: "15/05/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setup

source('config.R')
require('RPostgreSQL')
require('ggplot2')

drv <- dbDriver("PostgreSQL")
# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = pgdb,
                 host = pghost, port = pgport,
                 user = pguser, password = pgpassword)
```

## Overview

The purpose of this report is to gain insight on the behaviour of users in the first week of their trial.

To gain some insight into how first week product usage impacts customer retetion, the number of logins a user performs in their first week was compared to users that purchased a license compared to those who didn't

The following query is run to gain the dataset.

```{sql, connection=con, output.var="logincount"}
select
    u.user_id
  , u.active
  , count ( e.* ) as login_count
from
  only sc.sc_user u
  left outer join
  sc.sc_user_event e
    on u.user_id = e.user_id
    and e.event_ts >= u.signup_ts
    and e.event_ts <= u.signup_ts + interval '7 day'
    and e.event_type = 'login'
group by
    u.user_id
  , u.active
```

## Findings

The following two bar charts display a count of users by the number of times they login to the application in their first week, split by whether the user purchased a license for the software.


```{r pressure, echo=FALSE}
g <- ggplot(logincount, aes(login_count))
g + geom_bar() +
  facet_grid(active ~ .) +
  theme_minimal() +
  ggtitle("Login count by whether user purchased subscription") +
  xlab("# First week logins") +
  ylab("# Users")
```

As the graph shows, users who did not purchase a license logged into the application significantly less than those who extended their license.

This may be evidence that user retention beyond their trial period could be improved by giving a user incentives to login to the application more often in their first week.
